A Method for the spectroscopic inference of fundamental stellar parameters

### 2014-05-21 tex/paper.tex
Updated paper for SA review

\newcommand{\prob}{{\rm prob}}
\newcommand{\qN}{\{q_i\}_{i=1}^N}
\newcommand{\qM}{\{q_{im}\}_{i=1,m=0}^{N,M}}
\newcommand{\yN}{\{y_i\}_{i=1}^N}
\newcommand{\vt}{\vec{\theta}}
\newcommand{\vg}{\vt_{\star, {\rm grid}}}
\newcommand{\vpp}{\vt_{\star, {\rm post}}}
\newcommand{\vstar}{\vt_{\star}}
\newcommand{\finst}{f_{\lambda, {\rm inst}}}
\newcommand{\fsynth}{f_{\lambda, {\rm synth}}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vc}{\vec{c}}
\newcommand{\fM}{ \vec{{\bm M}}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ \vec{{\bm D}}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}

@@

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[width=0.85\textwidth]{figs/model_data.pdf}
  \caption{\textbf{Top}: Chebyshev polynomial which multiplies the model spectrum
   to account for inaccuracies in the flux-calibration. 
  \textbf{Middle}: The data spectrum and model spectrum, after it has been
   interpolated, post-processed, and multiplied by the Chebyshev polynomial.
  \textbf{Bottom}: Residuals from the model fit. Note the large residual at
   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{1}.}
\label{fig:model_data}
\end{center}
\end{figure*}

@@

Libraries of synthetic spectra are generally computed at high resolution ($R
 \geq $100,000), sampled with many pixels per resolution element, and do not
 include any rotational broadening or account for any instrumental effects. 
In order to transform a raw synthetic spectrum interpolated from the grid into
 one that matches the spectrum of a real star, we must post-process the spectrum
 to account for these secondary effects. 
In addition to these fundamental parameters, a star also has several observed
 properties that are a function of its kinematics, geometry, and location in our
 galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$,
 total extinction $A_V$, and solid angle $\Omega$. 
We call these secondary parameters $\vpp$, because we can model their effects
 on the synthetic spectra in a ``post-processing'' step by convolution with an
 appropriate kernel or multiplication with a smooth function. 
Together, we call $\vstar = \{\vg, \vpp\}$.

@@

A traditional way to deal with errors in the flux calibration is to avoid flux-calibration entirely and simply normalize both the data and the spectrum to the continuum level before comparison.  While this procedure may work well for hotter stars with well-defined continuua, normalizing spectra of cooler stars with large molecular features may introduce artificial features into the spectrum when the pseudo-continuum is incorrectly placed.  Instead, we choose to parameterize the residual flux calibration error by a low-order Chebyshev polynomial which is normalized to unity. The coefficients of the polynomial are parameters in the model that will be determined by the fitting algorithm. If we have multiple observations of spectrophotometric standards, then we can characterize the typical flux-calibration error of the spectrograph and put priors on the coefficients that will prevent overfitting of the polynomials.  For an example of the Chebyshev polynomial determined for a typical spectrum, see Figure~\ref{fig:model_data}. Using priors on the Chebyshev coefficients ensures that real spectral features, such as deep molecular bands in later type stars, are not accidentally treated as calibration artifacts. The overall normalization of the polynomial $c_0$ is degenerate with the solid angle of the star $\Omega$. Therefore we apply an additional constraint and enforce that the mean of the Chebyshev polynomial be equal to $1$. If there is only one order in the spectrum, such as with a long-slit spectrum or a single order of an echelle spectrum, then this means setting $c_0 = 1$. If we are fitting more than one order of an echelle spectrum then we adopt an independent Chebyshev polynomial for each order. This means that we chose the $c_0$ of a specific order to anchor to 1, while the $c_0$ for other orders are allowed to take on perturbations about $1$.

@@

\begin{figure*}[!htb]
\begin{center}
  \includegraphics{figs/badlines.pdf}
  \caption{A collection of spectral lines induced by imperfect model fits.
    From left to right: \textbf{Class 0} The majority of spectral lines
    ($\gtrsim 60$\%) will have minor differences in strength between data and
    model spectrum, which produce low-amplitude correlations in the residuals
    on the length scale of the width of a typical spectral line.  \textbf{Class
    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
    the model (in this case a line-blended \ion{Ca}{1}) leaves a large, highly correlated
    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
    of all lines), an extraneous line in the model leaves a large, highly
    correlated patch of positive residual.  \textbf{Class III}: If the line strengths are
    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
    correlated residuals of moderate amplitude.  The difficulty
    with class III lines is that for any specific line, there might exist a
    $\vstar$ that will fit the line, but there does not exist a $\vstar$ that
    will properly fit \emph{all} the lines.}
\label{fig:badlines}
\end{center}
\end{figure*}

However, the pixel residuals from a spectroscopic fit will often be correlated and will invalidate Equation~\ref{eqn:chi2}. These correlations arise because the CCD pixels of a spectrograph do not sample independent wavelengths of the spectrum. Instead, spectrographs are designed so that the LSF is properly sampled by at least a few pixels, meaning that adjacent pixels do not sample completely independent parts of the spectrum. Because of this, when a spectral feature differs between the model and the data it creates a residual which spans several adjacent data points. For a collection of correlated residuals produced by a typical spectroscopic fit, see Figure~\ref{fig:badlines}.

\begin{figure*}[!htb]
\begin{center}
  $\vcenter{\hbox{\includegraphics{figs/class0_residuals.pdf}}}$
  $\vcenter{\hbox{\includegraphics{figs/class0_autocorrelation.pdf}}}$
  \caption{\textbf{Left} The same residuals from Figure~\ref{fig:badlines}, panel 0, but shown at greater zoom to show the mildly covariant structure produced by slight mismatch between the data and model spectra.   
  \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\sim 3 -4$ pixels.}
\label{fig:class0}
\end{center}
\end{figure*}

There is an important distinction between astrophysical or instrumental ``noise'' and the amplitude and correlation length of the residuals from a spectroscopic fit. Noise introduced to a spectrum by astrophysical or instrumental effects is generally uncorrelated with wavelength. This is because the arrival of each photon to the detector is a discrete event, and while photons are scattered by the LSF, the particular scattering direction of each photon is independent from any other photon. We can verify this by measuring the autocorrelation of residuals in a region of well-fit stellar continuum, and show that it is flat. However, when we compare two spectra that have many absorption lines, even small differences between the data and model spectra will produce correlated \emph{residuals}. We can verify this correlation by estimating the autocorrelation of the residuals
\begin{equation}
  \hat{\rho}\,(k) = \frac{1}{(n - k)\,\sigma^2} \sum_{i = 1}^{n - k} (R_i - \mu_R)\,(R_{i+k} - \mu_R)
  \label{eqn:autocorrelation}
\end{equation}
over a region of spectrum of length $n$. Figure~\ref{fig:class0} shows a general residual structure induced by slight model mismatch and the autocorrelation of this series. There is significant correlation on length scales of $\sim 3 -4$ pixels, which is also the typical width of a spectral line for this spectrum.

Although the correlation in the residuals is artificially induced by an
imperfect model fit rather than originating from a physical source of
correlated noise, we can treat these two effects as mathematically similar. By
using a non-trivial covariance matrix with off-diagonal terms we can describe
the small-scale covariance between pixel residuals shown in
Figures~\ref{fig:badlines}~and~\ref{fig:class0}. A non-diagonal covariance also
requires that we use  a likelihood function (Equation~\ref{eqn:lnprob}) which
uses a matrix product instead of a sum over independent pixels.

We seek to understand and parameterize the structure of covariance in the
spectroscopic residuals, because ignoring this complexity will bias the estimates of
the model parameters $\vt$. As a simple analogy, consider the fit of a straight
line to a dataset.  If the noise in the dataset is correlated, then adjacent
data points might be offset from the linear trend in the same direction by a
similar amount.  If the covariant structure of the noise is ignored, a simple
$\chi^2$ will treat these correlated offsets as part of the linear trend, which
will result in a biased determination of the slope and intercept of the line,
typically with uncertainties that are too small. We face a mathematically similar problem with 
correlated pixel residuals from a spectroscopic fit.

We parameterize the covariance structure using covariance kernels, taken from
the canon of Gaussian processes. Covariance kernels describe the covariance
between two pixels $\lambda_i$ and $\lambda_j$.  These kernels are analogous to
the cosmological two-point correlation function, where the distance
between two galaxies is used instead of wavelength of light.

\paragraph{Global and mildly covariant structure}
To account for the mild covariance structure shown in Figure~\ref{fig:class0},
which is generally low amplitude and has a correlation length of only a few
pixels but is present throughout the majority of the spectrum, we use a
\emph{stationary} covariance kernel. Stationary covariance kernels, also called \emph{radial basis functions}, satisfy the property that the
amount of correlation only depends on the distance between the two pixels $r$.  For spectra,
we map $r$ to the velocity difference between two pixels
\begin{equation}
  r\,(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |
\end{equation}
Then the covariance kernel describes the expected covariance between pixel residuals
\begin{equation}
  k(\lambda_i, \lambda_j) =  \langle R_i \; R_j \rangle
  \label{eqn:expectation}
\end{equation}
In theory, we could chose from a variety of covariance kernels to parameterize this mildly covariant structure, as long as they satisfy the property that the degree of covariance smoothly declines with increasing radial distance. In practice, there are a few covariance kernels which are well-respected within the machine-learning community and frequently employed to parameterize covariance of this nature \citep{rw05}. We choose to use the \matern\ $\nu = 3/2$ kernel because it is one of the most commonly used kernels for this purpose and has an extensive history of use (cite?).
\begin{equation}
  k_{\nu = 3/2}(r,\, a,\, l) = a \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp 
   \left (- \frac{\sqrt{3} r}{l} \right )
\end{equation}
This kernel is an exponential tapered by a polynomial, which makes for a smooth transition from small scale covariance to negligible covariance at longer scales. In order to ensure sparsity of the covariance matrix for computational efficiency, we
 taper the \matern\ kernel for compact support using a Hann window
\begin{equation}
  w\,(r,\, r_0) = \left \{ 
    \begin{array}{cc}
    0.5 + 0.5 \cos (\pi r/ r_0 ) & r \le r_0 \\
    0 & r > r_0\\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
for a final covariance kernel of 
\begin{equation}
  k_{\rm global} (r,\, a,\, l,\, r_0) = w(r,\, r_0)\; k_{\nu = 3/2}(r,\, a,\, l) 
  \label{eqn:global}
\end{equation}
We add this covariance kernel to the independent pixel noise $\delta_{ij} \sigma_{ij}$ already present along
 the matrix diagonal, and introduce an additional parameter to allow a constant rescaling of the independent noise. Thus, each element of the covariance matrix is completely described by
\begin{equation}
  k(\lambda_i, \lambda_j |\, r,\, a,\, l,\, r_0,\, b) = k_{\rm global}(r, a, l, r_0) + b\,\delta_{ij} \sigma_{ij}
\end{equation}
Therefore, the \emph{hyperparameters} of this covariance kernel directly dictate the structure of the covariance matrix. By varying the hyperparameters, we can create a covariance matrix which accurately mimics the correlated structure of the pixel residuals. Most importantly, when the hyperparameters are correctly chosen to properly model this correlated structure, the \emph{likelihood function will be maximized} (Equation~\ref{eqn:prob}) while encapsulating the \emph{true uncertainty in the model-data fit}.

\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/matern_matrix.pdf}
\includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global}. There is a small degree of covariance in elements within a few pixels off the diagonal, which quickly tapers off such that the majority of the matrix remains sparse ($\sigma_{ij} = 0$). This matrix can be used to model spectrum which is correlated on a $\sim$few pixel scale.
\textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix. Top: the same residuals shown in Figure~\ref{fig:class0}, plotted against three sets of simulated residuals. The amplitude and correlation length of the simulated residuals closely approximates the structure of the actual pixel residuals.}
\label{fig:matern}
\end{center}
\end{figure*}

The left panel of Figure~\ref{fig:matern} shows what the covariance matrix might look like for reasonable choices of the kernel hyperparameters. To verify that this covariance matrix indeed mimics the correlated structure seen in the residuals, we can draw a random vector of residuals $R_{\rm rand}$ from our noise model and compare it to the actual pixel residuals. Because we are using a multidimensional Gaussian as our likelihood function (Equation~\ref{eqn:prob}), we have assumed that the underlying residual structure in a spectroscopic fit is described by a multidimensional Gaussian. This is directly analogous to the implicit assumption of independent Gaussian noise when using a standard $\chi^2$ likelihood (Equation~\ref{eqn:chi2}). If one wished to compare the residuals that resulted from a $\chi^2$ fit to the noise in the data set, then one could simulate $R_{\rm rand}$ by drawing $N$ random samples from a univariate normal distribution with zero mean and variance $\sigma_i^2$. We can make the same test with our multidimensional Gaussian likelihood function, however in this case the mean is a vector $\vec{\mu} = 0$ with length $N$ and the variance is described by the covariance matrix $C$. The residual vector is then a random draw from this multivariate normal distribution 
\begin{equation}
  R_{\rm rand} \sim {\cal N}\left( \vec{\mu}=0,\, C\right)
\end{equation}
Functions to generate random samples from a multidimensional Gaussian are implemented in most numerical programming languages.

\paragraph{Specific line covariance} Even after the mildly covariant structure has been accounted for by the global covariance kernel (Equation~\ref{eqn:global}), there still exist a significant number large amplitude pixel residuals that are \emph{highly correlated}. These patches of large amplitude residuals are caused by incorrect lines like those in the class I, II, and III examples in Figure~\ref{fig:badlines}. To parameterize the covariance in a given patch of these residuals, we introduce an additional \emph{non-stationary} covariance kernel. In this context, non-stationary means that the degree of covariance explicitly depends on the values of $\lambda_i$ and $\lambda_j$, and not simply the distance between them, therefore we can now individually target the large residuals induced by the mismatch of a \emph{specific} spectral line.

Because the shape of a spectral line can be generally approximated by a Gaussian function, and the difference of two Gaussians with the same mean is also another Gaussian--the same shape of residuals that are shown in Figure~\ref{fig:badlines}--we use Equation~\ref{eqn:expectation} to design a new covariance kernel about the Gaussian function. If we expect the shape of the residuals to be a Gaussian function about the line center $\lambda_\mu$
\begin{equation}
  R(\lambda |\, a, \lambda_\mu, \sigma) = \frac{a}{\sqrt{2 \pi} \sigma} \exp \left (-\; \frac{r^2(\lambda, \lambda_\mu)}{2 \sigma^2} \right )
\end{equation}
then the Gaussian covariance kernel is
\begin{align}
k_{\rm G}(\lambda_i, \lambda_j |\, a, \lambda_\mu, \sigma) &= \langle R(\lambda_i)\; R(\lambda_j) \rangle \\
&= \frac{a^2}{2 \pi \sigma} \exp \left ( - \, \frac{r^2(\lambda_i, \lambda_\mu) 
    + r^2(\lambda_j, \lambda_\mu)}{2 \sigma^2}\right )
\end{align}
We again enforce compact support by tapering the kernel with a Hann window (Equation~\ref{eqn:Hann}), and so the covariance kernel for the highly correlated residuals induced by a poor line fit becomes
\begin{equation}
  k_{\rm line}(\lambda_i, \lambda_j |\, a, \lambda_\mu, \sigma, r, r_0) = w(r,\, r_0)\, k_{\rm G}(\lambda_i, \lambda_j |\, a, \lambda_\mu, \sigma)
  \label{eqn:line}
\end{equation}

\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/matrix_region_covariance.png}
\includegraphics[width=0.4\textwidth]{figs/random_draw.png}
\caption{\textbf{Left} a covariance matrix generated with the region kernel.
\textbf{Right} general spectroscopic residuals near a mismatched stellar line
 overlaid with a sample draw using the covariance matrix, showing that the two
 look similar in structure and amplitude. 
\protect \todo{Similar improvements to
  Figure~\ref{fig:matern}.}}
\label{fig:region}
\end{center}
\end{figure*}

\todo{Similar paragraph about visualizing the covariance matrix, drawing
  samples. References to Figure~\ref{fig:region}.}

non-stationary kernel localizes increased variance to a specific
    spectral line

better than just masking out the region, since these contain info

>>>>

\newcommand{\prob}{{\rm prob}}
\newcommand{\qN}{\{q_i\}_{i=1}^N}
\newcommand{\qM}{\{q_{im}\}_{i=1,m=0}^{N,M}}
\newcommand{\yN}{\{y_i\}_{i=1}^N}

%\vt stands for ``vector theta''
\newcommand{\vt}{ {\bm \theta}}
\newcommand{\vg}{\vt_{\star, {\rm grid}}}
\newcommand{\vpp}{\vt_{\star, {\rm post}}}
\newcommand{\finst}{f_{\lambda, {\rm inst}}}
\newcommand{\fsynth}{f_{\lambda, {\rm synth}}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vtstar}{\vt_{\star}}
\newcommand{\vtcheb}{\vt_{\rm Cheb}}
\newcommand{\vtglobal}{\vt_{\rm global}}
\newcommand{\vtorder}[1]{\vt_{{\rm order}_{#1}}}
\newcommand{\vtorders}{\vt_{\rm orders}}
\newcommand{\vtline}[1]{\vt_{{\rm line}_{#1}}}
\newcommand{\vtlines}{\vt_{\rm lines}}
\newcommand{\vtcov}{\vt_{\rm cov}}
\newcommand{\fM}{ \vec{{\bm M}}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ \vec{{\bm D}}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}

@@

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[width=0.85\textwidth]{figs/model_data.pdf}
  \caption{\textbf{Top}: Chebyshev polynomial which multiplies the model spectrum
   to account for inaccuracies in the flux-calibration. 
  \textbf{Middle}: The data spectrum and model spectrum, after it has been
   interpolated, post-processed, and multiplied by the Chebyshev polynomial.
  \textbf{Bottom}: Residuals from the model fit. Note the large residual at
   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{2}.}
\label{fig:model_data}
\end{center}
\end{figure*}

@@

Libraries of synthetic spectra are generally computed at high resolution ($R
 \geq $100,000), sampled with many pixels per resolution element, and do not
 include any rotational broadening or account for any instrumental effects. 
In order to transform a raw synthetic spectrum interpolated from the grid into
 one that matches the spectrum of a real star, we must post-process the spectrum
 to account for these secondary effects. 
In addition to these fundamental parameters, a star also has several observed
 properties that are a function of its kinematics, geometry, and location in our
 galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$,
 total extinction $A_V$, and solid angle $\Omega$. 
We call these secondary parameters $\vpp$, because we can model their effects
 on the synthetic spectra in a ``post-processing'' step by convolution with an
 appropriate kernel or multiplication with a smooth function. 
Together, we call $\vtstar = \{\vg, \vpp\}$.

@@

A traditional way to deal with errors in the flux calibration is to avoid flux-calibration entirely and simply normalize both the data and the spectrum to the continuum level before comparison.  While this procedure may work well for hotter stars with well-defined continuua, normalizing spectra of cooler stars with large molecular features may introduce artificial features into the spectrum when the pseudo-continuum is incorrectly placed.  Instead, we choose to parameterize the residual flux calibration error by a low-order Chebyshev polynomial which is normalized to unity. The coefficients of the polynomial are parameters in the model that will be determined by the fitting algorithm, $\vtcheb = \{c_0, c_1, \ldots, c_N \}$. If we have multiple observations of spectrophotometric standards, then we can characterize the typical flux-calibration error of the spectrograph and put priors on the coefficients that will prevent overfitting of the polynomials.  For an example of the Chebyshev polynomial determined for a typical spectrum, see Figure~\ref{fig:model_data}. Using priors on the Chebyshev coefficients ensures that real spectral features, such as deep molecular bands in later type stars, are not accidentally treated as calibration artifacts. The overall normalization of the polynomial $c_0$ is degenerate with the solid angle of the star $\Omega$. Therefore we apply an additional constraint and enforce that the mean of the Chebyshev polynomial be equal to $1$. If there is only one order in the spectrum, such as with a long-slit spectrum or a single order of an echelle spectrum, then this means setting $c_0 = 1$. If we are fitting more than one order of an echelle spectrum then we adopt an independent Chebyshev polynomial for each order. This means that we chose the $c_0$ of a specific order to anchor to 1, while the $c_0$ for other orders are allowed to take on perturbations about $1$.

@@

\begin{figure*}[!htb]
\begin{center}
  \includegraphics{figs/badlines.pdf}
  \caption{A collection of spectral lines which have imperfect model fits.
    From left to right: \textbf{Class 0} The majority of spectral lines
    ($\gtrsim 60$\%) will have minor differences in strength between data and
    model spectrum, which produce low-amplitude correlations in the residuals
    on the length scale of the width of a typical spectral line.  \textbf{Class
    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
    of all lines), an extraneous line in the model leaves a large, highly
    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
    correlated residuals of moderate amplitude.  The difficulty
    with class III lines is that for any specific line, there might exist a
    $\vtstar$ that will fit the line, but there does not exist a $\vtstar$ that
    will properly fit \emph{all} the lines.}
\label{fig:badlines}
\end{center}
\end{figure*}

However, the pixel residuals from a spectroscopic fit will often be correlated and will invalidate Equation~\ref{eqn:chi2}. These correlations arise because the CCD pixels of a spectrograph do not sample independent wavelengths of the spectrum. Instead, spectrographs are designed so that the LSF is properly sampled by at least a few pixels, meaning that adjacent pixels do not sample completely independent parts of the spectrum. Because of this, when a spectral feature differs between the model and the data, it creates a residual which spans several adjacent data points. For a collection of correlated residuals produced by a typical spectroscopic fit, see Figure~\ref{fig:badlines}.

\begin{figure*}[!htb]
\begin{center}
  $\vcenter{\hbox{\includegraphics{figs/class0_residuals.pdf}}}$
  $\vcenter{\hbox{\includegraphics{figs/class0_autocorrelation.pdf}}}$
  \caption{\textbf{Left} The same residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.   
  \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
\label{fig:class0}
\end{center}
\end{figure*}

There is an important distinction between astrophysical or instrumental ``noise'' and the residuals from a spectroscopic fit. Noise introduced to a spectrum by astrophysical or instrumental effects is generally uncorrelated with wavelength. This is because the arrival of each photon to the detector is a discrete event, and while photons are scattered by the LSF, the particular scattering direction of each photon is independent from any other photon. We can verify this by measuring the autocorrelation of residuals in a region of well-fit stellar continuum, and show that it is flat. If the model spectrum is \emph{perfect}, then the residuals from a spectroscopic fit will be uncorrelated and equivalent to the astrophysical and instrumental noise introduced into the measurement. However, when we compare a data spectrum and a model spectrum that have many absorption lines, even small differences will produce correlated \emph{residuals}. We can verify this correlation by estimating the autocorrelation of the residuals as a function of pixel offset $k$
\begin{equation}
  \hat{\rho}\,(k) = \frac{1}{(n - k)\,\sigma^2} \sum_{i = 1}^{n - k} (R_i - \mu_R)\,(R_{i+k} - \mu_R)
  \label{eqn:autocorrelation}
\end{equation}
over a region of spectrum of length $n$ pixels. Figure~\ref{fig:class0} shows a general residual structure induced by slight model mismatch and the autocorrelation of this series. There is significant correlation for offsets of $\lesssim 4$ pixels, which is also the typical width of a spectral line for this spectrum.

Although the correlation in the residuals is artificially induced by an
imperfect model fit rather than originating from a physical source of
correlated noise, we can treat these two effects as mathematically similar. By
using a non-trivial covariance matrix with off-diagonal terms we can describe
the small-scale covariance between pixel residuals shown in
Figures~\ref{fig:badlines}~and~\ref{fig:class0}. A non-diagonal covariance also
requires that we use  a likelihood function  which uses a matrix product instead of a sum over independent pixels (Equation~\ref{eqn:lnprob}).

We seek to understand and parameterize the structure of covariance in the
spectroscopic residuals, because ignoring this complexity will bias the estimates of
the model parameters $\vt$. As a simple analogy, consider the fit of a straight
line to a dataset.  If the noise in the dataset is correlated, then nearby
data points might be offset from the linear trend in the same direction by a
similar amount.  If the covariant structure of the noise is ignored, a simple
$\chi^2$ will treat these correlated offsets as due to the linear trend, which
will result in a biased determination of the slope and intercept of the line,
typically with uncertainties that are too small. We face a mathematically similar problem with 
correlated pixel residuals from a spectroscopic fit.

We parameterize the covariance structure using covariance kernels, taken from
the field of Gaussian processes. Covariance kernels describe the covariance
between two pixels $\lambda_i$ and $\lambda_j$.  These kernels are analogous to
the cosmological two-point correlation function, where the distance
between two galaxies is used instead of the wavelength of light.

\paragraph{Global and mildly covariant structure}
To account for the mild covariance structure shown in Figure~\ref{fig:class0},
which is generally low amplitude and has a correlation length of only a few
pixels but is present throughout the majority of the spectrum, we use a
\emph{stationary} covariance kernel. Stationary covariance kernels, also called \emph{radial basis functions}, satisfy the property that the
amount of correlation only depends on the distance between the two pixels $r$.  For spectra,
we map $r$ to the velocity difference between two pixels
\begin{equation}
  r\,(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |
\end{equation}
Then the covariance kernel describes the expected covariance between pixel residuals
\begin{equation}
  k(\lambda_i, \lambda_j) =  \langle R_i \; R_j \rangle
  \label{eqn:expectation}
\end{equation}
In theory, we could chose from a plethora of covariance kernels to parameterize this mildly covariant structure, as long as they satisfy the property that the degree of covariance smoothly declines with increasing radial distance. In practice, there are a few covariance kernels which are well-respected within the machine-learning community and frequently employed to parameterize covariance of this nature \citep{rw05}. We choose to use the \matern\ $\nu = 3/2$ kernel because it is one of the most commonly used kernels for this purpose and has an extensive history of use (cite?). 
\begin{equation}
  k_{\nu = 3/2}(r\, |\, a_{\rm g},\, l) = a_{\rm g} \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp 
   \left (- \frac{\sqrt{3} r}{l} \right )
\end{equation}
This kernel is an exponential tapered by a polynomial, which makes for a smooth transition from small scale covariance to negligible covariance at longer scales. In order to ensure sparsity of the covariance matrix for computational efficiency, we
 taper the \matern\ kernel for compact support using a Hann window
\begin{equation}
  w\,(r |\, r_0) = \left \{ 
    \begin{array}{cc}
    0.5 + 0.5 \cos (\pi r/ r_0 ) & r \le r_0 \\
    0 & r > r_0\\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
for which we generally fix $r_0$ to be some fixed multiple of the length scale, say $r_0 = 4 l$. Then, if we define the global parameters $\vtglobal = \{a_g, l \}$, we have a final covariance kernel of 
\begin{equation}
  k_{\rm global} (r\,|\,\vtglobal) = w(r |\, r_0 = 4l)\; k_{\nu = 3/2}(r |\, a_{\rm g},\, l) 
  \label{eqn:global}
\end{equation}
\todo{This might be a good place to actually include a simple plot of $k_{\rm global}$ as a function of $r$, as long as we aren't providing too many figures. Maybe it could be combined with Figure~\ref{fig:matern}.}

We add this covariance kernel to the independent pixel noise $\delta_{ij} \sigma_{ij}$ already present along
 the matrix diagonal, and introduce an additional parameter to allow a constant rescaling of the independent noise. Then, each element of the covariance matrix is completely described by
\begin{equation}
  k(\lambda_i, \lambda_j |\vtglobal, \, b) = k_{\rm global}(r | \vtglobal) + b\,\delta_{ij} \sigma_{ij}
\end{equation}
Therefore, the \emph{hyperparameters} of this covariance kernel directly determine the structure of the covariance matrix. By varying the hyperparameters, we can create a covariance matrix which accurately mimics the correlated structure of the pixel residuals. Most importantly, when the hyperparameters are correctly chosen to properly model this correlated structure, the likelihood function will be maximized (Equation~\ref{eqn:prob}) while also encapsulating the true uncertainty in the spectroscopic fit.

\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/matern_matrix.pdf}
\includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global} and common values for the hyperparameters. There is a small degree of covariance in elements within a few pixels off the diagonal, which quickly tapers off so that the majority of the matrix remains sparse ($\sigma_{ij} = 0$). This matrix is used to model a spectrum which is correlated on a $\sim$few pixel scale.
\textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix. The top panel shows the same residuals shown in Figure~\ref{fig:class0}, and below are plotted three sets of simulated residuals. The amplitude and correlation length of the simulated residuals closely approximates the structure of the actual pixel residuals.}
\label{fig:matern}
\end{center}
\end{figure*}

The left panel of Figure~\ref{fig:matern} shows what the covariance matrix
might look like for reasonable choices of the kernel hyperparameters. To verify
that this covariance matrix indeed mimics the correlated structure seen in the
residuals, we can draw a random vector of residuals $R_{\rm rand}$ from our
noise model and compare it to the actual pixel residuals. Because we are using
a multidimensional Gaussian as our likelihood function
(Equation~\ref{eqn:prob}), we have assumed that the underlying residual
structure in a spectroscopic fit is described by a multidimensional Gaussian.
This is directly analogous to the implicit assumption of independent Gaussian
noise when using a standard $\chi^2$ likelihood (Equation~\ref{eqn:chi2}). If
one wished to compare the residuals that resulted from a $\chi^2$ fit to the
noise in the data set, then one could simulate $R_{\rm rand}$ by drawing $N$
random samples from a univariate normal distribution with zero mean and
variance $\sigma_i^2$. We can make the same test with our multidimensional
Gaussian likelihood function, however in this case the mean is a vector
$\vec{\mu} = 0$ with length $N$ and the variance is described by the covariance
matrix $C$. The residual vector is then a random draw from this multivariate
normal distribution 
\begin{equation}
  R_{\rm rand} \sim {\cal N}\left( \vec{\mu}=0,\, C\right)
\end{equation}
Functions to generate random samples from a multidimensional Gaussian are implemented in most numerical programming languages. As we see in the right panel of Figure~\ref{fig:matern}, the structure and amplitude of the fake residuals closely mimics that of the actual data residuals, suggesting that the covariance kernel faithfully reproduces the covariance induced by the data-model mismatch.

\paragraph{Specific line covariance} Even after the mildly covariant structure has been accounted for by the global covariance kernel (Equation~\ref{eqn:global}), there still exist a significant number large amplitude pixel residuals that are \emph{highly correlated}. These patches of large amplitude residuals are caused by incorrect synthetic lines like those in the class I, II, and III examples in Figure~\ref{fig:badlines}. To parameterize the covariance in a given patch of these residuals, we introduce an additional \emph{non-stationary} covariance kernel. In this context, non-stationary means that the degree of covariance explicitly depends on the values of $\lambda_i$ and $\lambda_j$ and not simply the distance between them, enabling us to individually target the large residuals induced by the mismatch of a \emph{specific} spectral line.
